---
title: "p8105_hw2_cz2750"
output: github_document
date: "2023-10-04"
---

## P1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(tidyverse)
library(readxl)
options(warn = -1)
```

#Clean the data in pols-month.csv
```{r}
pols_month = read.csv(file = "pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = month.name[as.integer(month)],
         president = ifelse(prez_dem == 1, "dem", ifelse(prez_gop == 1, "gop", NA))) %>%
  select(-prez_dem, -prez_gop, -day)

pols_month
```

#Clean the data in snp.csv
```{r}
snp = read.csv(file = "snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(year = ifelse(year >= 0 & year <= 15, paste("20", year, sep = ""), paste("19", year, sep = ""))) %>%
  mutate(month = month.name[as.integer(month)]) %>%
  arrange(year, month) %>%
  select(year, month, day, everything()) %>%
  select(-day)

snp
```
#Tidy the unemployment data
```{r}
unemployment = read.csv(file = "unemployment.csv") %>%
  pivot_longer(cols = -Year, names_to = "month", values_to = "unemployment") %>%
  mutate(month = tolower(month)) %>%
  mutate(month = case_when(
    month %in% c("jan") ~ "January",
    month %in% c("feb") ~ "February",
    month %in% c("mar") ~ "March",
    month %in% c("apr") ~ "April",
    month %in% c("may") ~ "May",
    month %in% c("jun") ~ "June",
    month %in% c("jul") ~ "July",
    month %in% c("aug") ~ "August",
    month %in% c("sep") ~ "September",
    month %in% c("oct") ~ "October",
    month %in% c("nov") ~ "November",
    month %in% c("dec") ~ "December",
    TRUE ~ NA_character_ )) %>%
  select(year=Year, month, unemployment, everything())

unemployment
```
#Merge the datasets
```{r}
result_data <- left_join(pols_month, snp, by = c("year", "month"))
unemployment$year <- as.character(unemployment$year)
BDATASET <- left_join(result_data, unemployment, by = c("year", "month"))

BDATASET
```
#Describe the resulting dataset

I named the dataset "BDATASET".The "Bdataset" is a merged dataset that combines information from three different sources: pols, snp, and unemployment. It contains data on political polling, stock market performance, and unemployment rates. Of these, "pols_month" provides political data." snp "provides financial data." unemployment "provides economic data in multiple formats.

For the "BDATASET":
Dimension: The dataset has a certain number of rows (observations) and columns (variables).
Range of Years: The dataset covers a range of years, from the earliest year to the latest year present in the original datasets.
Key Variables: The key variables in this dataset are year, month, president, snp_close, and unemployment_rate. year and month are used as keys for merging, president indicates the political party (GOP or DEM), snp_close represents the closing value of the S&P 500 index, and unemployment_rate shows the unemployment rate for a specific year and month.

The resulting dataset can be used for a variety of analyses, including exploring the relationship between political events, stock market performance, and unemployment over time.

## P2

#Read and Clean the Mr. Trash Wheel Dataset
```{r}
mr_trash_wheel = read_xlsx( 
  "202309 Trash Wheel Collection Data.xlsx",
  cell_cols("A:N"),
  sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(homes_powered_N = weight_tons*500/30,
         wheel_type = c("MT")) %>% 
  relocate(wheel_type,dumpster,year,month) 

mr_trash_wheel
```
#Read and Clean the Professor Trash Wheel Dataset
```{r}
professor_trash_wheel = read_xlsx( 
  "202309 Trash Wheel Collection Data.xlsx",
  cell_cols("A:N"),
  sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(homes_powered_N = weight_tons*500/30,
         wheel_type = c("MT")) %>% 
  relocate(wheel_type,dumpster,year,month) 

professor_trash_wheel
```


#Read and Clean the Gwynnda Dataset
```{r}
Gwynnda_trash_wheel = read_xlsx( 
  "202309 Trash Wheel Collection Data.xlsx",
  cell_cols("A:N"),
  sheet = "Gwynnda Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(homes_powered_N = weight_tons*500/30,
         wheel_type = c("MT")) %>% 
  relocate(wheel_type,dumpster,year,month) 

Gwynnda_trash_wheel
```
#Combine Datasets
```{r}
mr_trash_wheel$year <- as.double(mr_trash_wheel$year)
combined_dataset <- bind_rows(mr_trash_wheel, 
                         professor_trash_wheel, 
                         Gwynnda_trash_wheel)

combined_dataset
```
#Describe the Data

I named the dataset "combined_dataset".The dataset "combined_dataset" is a combined dataset that includes data from Mr. Trash Wheel, Professor Trash Wheel and Gwynnda.There are a total of `r nrow(combined_dataset)` observations in the combined dataset, representing various garbage collection events.

Key Variables:
date: The date of data collection.
total_weight: The total weight of trash collected.
trash_count: The count of trash items collected.
cigarette_butts: The count of cigarette butts collected.
homes_powered: The number of homes powered, if applicable.
trash_wheel: An identifier for the Trash Wheel (Mr. Trash Wheel, Professor Trash Wheel, or Gwynnda).

The total weight of garbage collected by Professor Garbage Truck is `r sum(professor_trash_wheel$weight_tons)` tons, and the total number of cigarette butts collected by Gwenda in July 2021 is `r format(sum(Gwynnda_trash_wheel$cigarette_butts[1:5]), scientific = F)`.


## P3

#Import, Clean, and Tidy the Baseline Demographics Dataset
```{r}
baseline_data = read.csv(file = "MCI_baseline.csv", skip=1) %>%
  janitor::clean_names() %>%
  mutate(sex = factor(sex, labels = c("Female", "Male")),        
         apoe4 = factor(apoe4, labels = c("APOE4 non-carrier", "APOE4 carrier")),        
         age_at_onset = as.numeric(age_at_onset)) %>%
  drop_na(age_at_onset)

baseline_data
```
#Describe the Data

Important steps in the import process and related functions of the dataset: we use the `janitor` function to clean up the column names, `mutate` to code the categorical variables, to convert the data types if necessary, and to deal with missing data in preparation for further analysis of the dataset.

How many participants were recruited: r nrow(baseline_data)
How many of them have MCI: r sum(!is.na(baseline_data$age_at_onset))
Average baseline age: r round(mean(baseline_data$current_age, na.rm=TRUE),digits = 0)
Proportion of female APOE4 carriers in the study: r round(sum(baseline_data$sex == "Female" & baseline$apoe4 == "APOE4 carrier") / sum(baseline_data$sex == "Female"), digits = 4) * 100%

#Import, Clean, and Tidy the amyloid Demographics Dataset
```{r}
amyloid_data = read.csv(file = "mci_amyloid.csv", skip=1) %>%
  janitor::clean_names() %>%
  pivot_longer(baseline:time_8,
               names_prefix="time_",
               names_to="years_since_diagnosed",
               values_to="apoe4") %>%
  mutate(years_since_diagnosed=replace(years_since_diagnosed, years_since_diagnosed=="baseline", 0)) %>%
  mutate(years_since_diagnosed=as.numeric(years_since_diagnosed))

amyloid_data
```
#Describe the Data

Important steps in the import process and related functions of the dataset: We use `read_csv` to read the CSV file and skip the first line. Next, column names are normalized and cleaned using the `janitor::clean_names()` function. Categorized variables were coded for analysis and data types were converted as needed.

How many participants were recruited: r nrow(distinct(amyloid_data, study_id))

#Check whether certain participants appear only in the baseline or amyloid dataset.
```{r}
baseline <- baseline_data %>%
  anti_join(amyloid_data, by = c("id" = "study_id")) %>%
  select(id)

amyloid <- amyloid_data %>%
  anti_join(baseline_data, by = c("study_id" = "id")) %>%
  select(study_id)
```

#Describe

Participants unique to the baseline dataset:
r nrow(baseline)
Participants unique to the amyloid dataset: 
r nrow(amyloid)

#Combine Datasets
```{r}
ADATASET <- baseline_data %>%
  inner_join(amyloid_data, by = c("id" = "study_id"))
```

#Describe the Data

I named the dataset "ADATASET".The"ADATASET" retains participants that appear in both the "baseline" and "amyloid" datasets.

#Export the result as a CSV
```{r}
write.csv(ADATASET, "ADATASET.csv", row.names = FALSE)
```

